Ore Polynomials in Sage

### 2013-06-10 paper/main.tex

-- options for handling singularities

-- rectangular splitting

-- available for shift algebras.

>>>>

-- available for shift and q-shift algebras.

We compute an annihilator for the sequence
$c(n) = \sum_{k=0}^n 1 / k!$:

\begin{sageexample}
  sage: R.<n> = ZZ[]; RS.<Sn> = OreAlgebra(R, 'Sn')
  sage: inverse_factorials = (n + 1) * Sn - 1
  sage: partial_sums = inverse_factorials.annihilator_of_sum()
  sage: partial_sums
\end{sageexample}

The \verb|to_list| method returns the first few values of a sequence,
given the initial values:

\begin{sageexample}
  sage: L = partial_sums.to_list([1, 2], 8)
  sage: L
  sage: N(L[7])
\end{sageexample}

In some cases (for example when the base ring is $\mathbb{Z}$ or
$\mathbb{Z}[x]$), isolated values can be
computed asymptotically faster for large $n$ using the binary splitting
technique.
The \verb|forward_matrix_bsplit| method, called with argument $n$,
returns a matrix $P$ and a denominator $Q$ such that $P / Q$ multiplied a
column vector of initial values $c_0, c_1, \ldots$
yields $c_n, c_{n+1}, \ldots$.
This way, computing $10^5$ digits of $e$ takes a fraction of a second:

\begin{sageexample}
  sage: e_approx = N(e, 400000)
  sage: P, Q = partial_sums.forward_matrix_bsplit(25600)
  sage: u = Matrix([[e_approx], [e_approx]]) - P * Matrix([[1], [2]]) / Q
  sage: u.change_ring(RealField(20))
\end{sageexample}
